---
bibliography: ../rforhta.bib
link-citations: true
---

## Data Visualisation

```{r}
library(here)
library(tidyverse)
source(here("_common.R"))
```

The following data used is the Innovation Scorecard (NICE Technology
Appraisals in the NHS in England) to October 2024 at the ICB level. If
you're not familiar with this dataset, it is found here (@zotero-739)
and "aims to improve transparency within the NHS of what treatments
recommended by NICE are available at a local level. The local level can
be within trusts and ICBs as well as national and NHS England region
levels. The Innovation Scorecard has been published with the intention
of assisting the NHS in the identification of variation which, through
discussion, can be explained, challenged, or acted upon. It is not
intended to be used for performance management.".

```{r}
dat_grp <- read.csv(file = here("./data/nice-tech-apps-eng-oct24-ICB-Group.csv"))

dat_grp1 <- dat_grp %>% filter(year == "2023_24", quarter == 1, data_source == "Secondary Care", treatment_type == "medicine")

```

```{r}
head(dat_grp1)
unique(dat_grp1$treatment_name)
```

```{r}
dat_ut <- read.csv(file = here("./data/nice-tech-apps-eng-oct24-ICB-Utilisation.csv"))

dat_ut1 <- dat_grp %>% filter(year == "2023_24", quarter == 1, data_source == "Secondary Care", treatment_type == "medicine")
```

```{r}
head(dat_ut1)
unique(dat_ut1$treatment_name)
```


```{r}
dat1 %>% filter(high_level_condition == "Breast Cancer") %>%  ggplot(aes(x = treatment_name, y = value)) +
  geom_boxplot(fill = "lightblue", alpha = 0.7) +
  theme_minimal() +
  labs(title = "Distribution of Utilisation by Treatment Type within Breast Cancer",
       y = "DDD per 100,000 population",
       x = "Treatment Type") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
dat1 %>%
  group_by(provider_name) %>%
  summarize(avg_utilization = mean(value)) %>%
  arrange(desc(avg_utilization)) %>%
  head(10) %>%
  ggplot(aes(x = reorder(provider_name, avg_utilization), y = avg_utilization)) +
  geom_point() +
  coord_flip() +
  theme_minimal() +
  labs(title = "Top 10 Providers by Average Utilization",
       x = "Provider Name",
       y = "Average DDD per 100,000 population")
```

```{r}
ggplot(dat1, aes(x = high_level_condition, y = value, fill = treatment_type)) +
  geom_violin(alpha = 0.7) +
  theme_minimal() +
  labs(title = "Treatment Utilization by Condition",
       y = "DDD per 100,000 population",
       x = "Condition") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  scale_y_continuous(labels = scales::comma)
```
