## Data Visualisation

```{r}
library(palmerpenguins)
library(ggthemes)
library(here)
library(tidyverse)
```

do penguins with longer flippers weigh more or less than pengiuns with
shorter flippers?

to make this easier, we define: - a variable is a quantity, quality or
property that you can measure - a value is the state of a variable when
you measure it, which may change from observation to observation - an
observation is a set of measurements made under similar conditions, an
observation will contain several values, each ass

```{r}
dat <- read.csv(file = here("./data/nice-tech-apps-eng-apr24-ICB-Utilisation.csv")) 

dat1 <- dat %>% filter(year == "2023_24", quarter == 1, data_source == "Secondary Care", data_type == "ICB Utilisation", treatment_type == "medicine")

dat_grp <- read.csv(file = here("./data//nice-tech-apps-eng-oct23-ICB-Group.csv"))

dat_grp1 <- dat_grp %>% filter(year == "2023_24", quarter == 1)
# data_source == "Secondary Care", data_type == "ICB Utilisation", treatment_type == "medicine"

```

```{r}
head(dat1)
```

```{r}
head(dat_grp1)
```

some changes